---
title: "Simple simulation"
date: "`r date()`"
---

```{r setup, include=FALSE}
library(lostruct)
library(colorspace)
library(jsonlite)
library(RColorBrewer)
fig.dim <- 4
knitr::opts_chunk$set(fig.width=2*fig.dim,fig.height=fig.dim,fig.align='center')
# set do.pdfs to TRUE to output PDFs of figures as well
if (!exists("do.pdfs")) { do.pdfs <- FALSE }
```
```{r plot_setup, include=FALSE}
layout_heights <- function (k,dl=0,ncol=1) {
    # to set up layout without 'dl' lines between plots
    # use like layout(1:5,heights=layout_heights(5))
    if (k==1) return(1)
    layout(matrix(seq_len(k*ncol),ncol=ncol))  # this changes par("csi")
    ds <- dl*par("lheight")*par("csi")
    eps=par("mai")[c(1,3)]
    dh=(par("din")[2]-sum(eps)-(k-1)*ds)/k
    return(c(eps[2]+dh+ds/2,rep(dh+ds,k-2),eps[1]+dh+ds/2)/par("din")[2])
}
pdf_copy <- function (
                width=6,
                height=width*knitr::opts_current$get("fig.height")/knitr::opts_current$get("fig.width"),
                plot.id=NULL,
                filename
                ) {
    if (missing(filename)) {
        file.id <- if (is.null(plot.id)) { "" } else { paste0("_",plot.id) }
        filename <- knitr::fig_path(paste(file.id,".pdf",sep=""))
    }
    cat("pdf version at:",filename)
    dev.print( file=filename, device=pdf, 
            width=width, height=height,
            pointsize=10,
            family="sans")
}
```

**Note:** to copy the resulting figures to the `writeup/` directory:
```
for x in figure/simple_example/*pdf; do mv $x "../../writeup/simple_example_$(basename ${x%-1.pdf}.pdf)"; done
```


Let's suppose there are three populations,
and on the first quarter of the chromosome,
the first two are close to each other;
on the middle hald, the last two are close,
and on the last quarter the first and the last are.

```{r simple_data, cache=TRUE}
base_covmat <- matrix(c(0.75, 0.25, 0.0,
                        0.25, 0.75, 0.0,
                        0.0, 0.0, 1.0), nrow=3)
pop_covmats <- list(base_covmat, 
                    base_covmat[3:1, 3:1], 
                    base_covmat[c(1,3,2), c(1,3,2)])
struct_breakpoints <- c(0.0, 0.25, 0.5, 0.75, 1.0)
struct_indices <- c(1,2,2,3)
nsamples <- c(50, 50, 50)
# ind_covmat <- diag(sum(nsamples))
# for (k in seq_along(nsamples)) {
#     ii <- sum(nsamples[seq_len(k-1)]) + 1:nsamples[k] 
#     ind_covmat[ii, ii] <- rexp(nsamples[k]^2, 100)
#     ind_covmat[ii, ii] <- ind_covmat[ii, ii] + t(ind_covmat[ii, ii])
# }
ind_covmat <- 0
pop_chols <- lapply(lapply(pop_covmats, chol),
                    function (x) { 
                        (x[rep(1:3, nsamples), rep(1:3, nsamples)] / sqrt(rep(nsamples, nsamples))
                         + ind_covmat)
                    })
nwindows <- 99
nloci <- 400
win.fn <- function (n, ...) {
    geno <- matrix(rnorm(sum(nsamples) * nloci), nrow=nloci)
    cholmat <- pop_chols[[struct_indices[findInterval(n/nwindows, struct_breakpoints, rightmost.closed=TRUE)]]]
    return( geno %*% cholmat )
}
attr(win.fn, "max.n") <- nwindows
attr(win.fn,"samples") <- paste0(rep(letters[seq_along(nsamples)], nsamples),
                                 unlist(lapply(nsamples, seq_len)))
attr(win.fn,"region") <- function (n) { 
    data.frame( chrom="matrix", start=((n-1)*nloci+1), end=n*nloci ) }
all.pcas <- eigen_windows(win.fn, k=2, do.windows=1:nwindows)
regions <- region(win.fn)()
all.lengths <- c("chrom1"=nwindows)
pc.distmat <- pc_dist( all.pcas, npc=2 )
na.inds <- is.na( all.pcas[,1] ) # there may be windows with missing data
mds.coords <- cbind( data.frame( 
                        chrom=rep(names(all.lengths),all.lengths),
                        window=unlist(lapply(all.lengths,seq_len)) ),
                        cmdscale( pc.distmat[!na.inds,!na.inds], k=2 )[ ifelse( na.inds, NA, cumsum(!na.inds) ), ] )
mds.cols <- (1:ncol(mds.coords))[-(1:2)]
```

OK, now we plot them:
```{r plotting_setup, echo=FALSE}
chroms <- unique(regions$chrom)
chrom.starts <- tapply( regions$start, regions$chrom, min, na.rm=TRUE )
chrom.ends <- tapply( regions$end, regions$chrom, max, na.rm=TRUE )
chrom.spacing <- floor(.05*mean(chrom.ends))
chrom.offsets <- c(0,cumsum(chrom.spacing+chrom.ends))
names(chrom.offsets) <- c(chroms,"end")
chrom.dividers <- c(0,chrom.offsets[-1])-chrom.spacing/2
chrom.mids <- chrom.dividers[-1] - diff(chrom.dividers)/2

# this is where to plot windows at when plotting with all chromosomes
regions$pos <- chrom.offsets[match(regions$chrom,chroms)]+(regions$start+regions$end)/2

chrom.cols <- rainbow_hcl(length(chroms), c=90, end=.9*360)[as.numeric(regions$chrom)]

#' Set up to plot all chromosomes together
#' will plot the vector of values 'y' along all the chromosomes.
chrom.plot <- function (y,ylab='',main='',chrom.labels=TRUE,...) {
    plot(0, type='n', xlim=range(chrom.offsets/1e6), ylim=range(y,finite=TRUE), 
         xlab='', xaxt='n', yaxt='n', ylab=ylab, main=main)
    if (length(chroms)>1) for (k in 1:floor(length(chroms)/2)) {
        rect( xleft=chrom.dividers[2*k-1]/1e6, xright=chrom.dividers[2*k]/1e6, 
             ybottom=par("usr")[3], ytop=par("usr")[4], 
             border=NA, col=adjustcolor("grey",0.25) )
    }
    abline( v=chrom.dividers/1e6, lty=3, col=adjustcolor("grey",0.5), lwd=2 )
    if (chrom.labels) axis( 1, at=chrom.mids/1e6, labels=paste("chromosome", chroms), las=0, tick=FALSE )
    points( regions$pos/1e6, y, ...)
}

pairs( mds.coords[,mds.cols], pch=20, col=adjustcolor(chrom.cols,0.75) )

plot_everything <- function () {
    mds.corners <- corners( mds.coords[,mds.cols[1:2]], prop=.05 )
    corner.cols <- brewer.pal(3,"Dark2")
    corner.pch <- c(15,17,19)
    ccols <- rep("black",nrow(mds.coords))
    cpch <- rep(20,nrow(mds.coords))
    for (k in 1:ncol(mds.corners)) {
        ccols[ mds.corners[,k] ] <- corner.cols[k]
        cpch[ mds.corners[,k] ] <- corner.pch[k]
    }
    # centroids of the corners in MDS space
    corner.mds <- do.call(rbind, lapply(1:ncol(mds.corners), 
                    function (ii){
                        colMeans(mds.coords[mds.corners[,ii],-(1:2)])
                    } ) )

    spacing <- 1
    opar <- par(mar=c(4,4,2,1)+.1,mgp=c(2.5,0.8,0))
    layout(matrix(c(rep(1,length(mds.cols)),1+seq_along(mds.cols)),ncol=2),
            widths=c(1,2), heights=layout_heights(length(mds.cols),dl=spacing,ncol=2))
    plot( mds.coords[,mds.cols[1:2]], pch=cpch, 
         col=adjustcolor(ccols,0.75),  asp=1,
         xaxt='n', yaxt='n',
         xlab="MDS coordinate 1", ylab="MDS coordinate 2" )
    points( corner.mds, pch=20, cex=5,
            col=adjustcolor(corner.cols,0.25))
    text( corner.mds, labels=seq_len(nrow(corner.mds)), 
            col=corner.cols, cex=2, lwd=2 )
    opar2 <- par(mar=c(par("mar"),spacing/2)[c(5,2,3,4)])
    for (k in mds.cols) {
        lastone <- (k==mds.cols[length(mds.cols)])
        if (lastone) { par(mar=c(par("mar"),opar2$mar[1])[c(5,2,3,4)]) }
        chrom.plot( mds.coords[,k], pch=20, 
                xlab=if (lastone) { "Position (Mb)"} else { "" }, # main=paste("MDS coordinate",match(k,mds.cols)),
                chrom.labels=lastone,
                ylab=colnames(mds.coords)[k],
                col=adjustcolor(ccols,0.75) )
        # do this for all but first
        par(mar=c(par("mar"),spacing/2)[c(1,2,5,4)])
    }
    par(opar)
    pdf_copy()
}
```
```{r base_case}
plot_everything()
```

## With missing data

Now we'll do the same thing, but with 50% missing data
on the first half of the chromosome.
```{r missing_data, cache=TRUE, depends="simple_data"}
win.fn <- function (n, ...) {
    geno <- matrix(rnorm(sum(nsamples) * nloci), nrow=nloci)
    cholmat <- pop_chols[[struct_indices[findInterval(n/nwindows, struct_breakpoints, rightmost.closed=TRUE)]]]
    out <- geno %*% cholmat
    if (n < nwindows/2) {
        out[rbinom(length(out), 1, 0.5)==1] <- NA
    }
    return(out)
}
attr(win.fn, "max.n") <- nwindows
attr(win.fn,"samples") <- paste0(rep(letters[seq_along(nsamples)], nsamples),
                                 unlist(lapply(nsamples, seq_len)))
attr(win.fn,"region") <- function (n) { 
    data.frame( chrom="matrix", start=((n-1)*nloci+1), end=n*nloci ) }
all.pcas <- eigen_windows(win.fn, k=2, do.windows=1:nwindows)
pc.distmat <- pc_dist( all.pcas, npc=2 )
na.inds <- is.na( all.pcas[,1] ) # there may be windows with missing data
mds.coords <- cbind( data.frame( 
                        chrom=rep(names(all.lengths),all.lengths),
                        window=unlist(lapply(all.lengths,seq_len)) ),
                        cmdscale( pc.distmat[!na.inds,!na.inds], k=2 )[ ifelse( na.inds, NA, cumsum(!na.inds) ), ] )
mds.cols <- (1:ncol(mds.coords))[-(1:2)]
plot_everything()
```

## With very different sample sizes

Now we'll do the same thing, but with very different sample sizes.
```{r sample_diffs, cache=TRUE, depends="simple_data"}
nsamples <- c(900, 10, 10)
ind_covmat <- diag(sum(nsamples))
for (k in seq_along(nsamples)) {
    ii <- sum(nsamples[seq_len(k-1)]) + 1:nsamples[k] 
    ind_covmat[ii, ii] <- rexp(nsamples[k]^2, 15)
    ind_covmat[ii, ii] <- ind_covmat[ii, ii] + t(ind_covmat[ii, ii])
}
pop_chols <- lapply(lapply(pop_covmats, chol),
                    function (x) { 
                        (x[rep(1:3, nsamples), rep(1:3, nsamples)] / sqrt(rep(nsamples, nsamples))
                         + ind_covmat)
                    })
nwindows <- 99
nloci <- 400
win.fn <- function (n, ...) {
    geno <- matrix(rnorm(sum(nsamples) * nloci), nrow=nloci)
    cholmat <- pop_chols[[struct_indices[findInterval(n/nwindows, struct_breakpoints, rightmost.closed=TRUE)]]]
    return( geno %*% cholmat )
}
attr(win.fn, "max.n") <- nwindows
attr(win.fn,"samples") <- paste0(rep(letters[seq_along(nsamples)], nsamples),
                                 unlist(lapply(nsamples, seq_len)))
attr(win.fn,"region") <- function (n) { 
    data.frame( chrom="matrix", start=((n-1)*nloci+1), end=n*nloci ) }
all.pcas <- eigen_windows(win.fn, k=2, do.windows=1:nwindows)
all.lengths <- c("chrom1"=nwindows)
pc.distmat <- pc_dist( all.pcas, npc=2 )
na.inds <- is.na( all.pcas[,1] ) # there may be windows with missing data
mds.coords <- cbind( data.frame( 
                        chrom=rep(names(all.lengths),all.lengths),
                        window=unlist(lapply(all.lengths,seq_len)) ),
                        cmdscale( pc.distmat[!na.inds,!na.inds], k=2 )[ ifelse( na.inds, NA, cumsum(!na.inds) ), ] )
mds.cols <- (1:ncol(mds.coords))[-(1:2)]

plot_everything()
```

And, *reweighted*:
```{r reweighted, cache=TRUE, depends="sample_diffs"}
all.pcas <- eigen_windows(win.fn, k=2, do.windows=1:nwindows, w=1/rep(sqrt(nsamples),nsamples))
all.lengths <- c("chrom1"=nwindows)
pc.distmat <- pc_dist( all.pcas, npc=2, w=1/rep(sqrt(nsamples),nsamples))
na.inds <- is.na( all.pcas[,1] ) # there may be windows with missing data
mds.coords <- cbind( data.frame( 
                        chrom=rep(names(all.lengths),all.lengths),
                        window=unlist(lapply(all.lengths,seq_len)) ),
                        cmdscale( pc.distmat[!na.inds,!na.inds], k=2 )[ ifelse( na.inds, NA, cumsum(!na.inds) ), ] )
mds.cols <- (1:ncol(mds.coords))[-(1:2)]

plot_everything()
```


